Empirical normative theories

### 2020-06-22 simple.tex
Fals single-try model (very messy)

\begin{document}
%\title{Theories of deliberated preference}
\title{Testable normative theories}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
	\href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

@@

Given any $\ar_0 \in \allargs, \phi \in \Phi$, define $\ar_0 \ileadstost[i] \phi$ iff $\forall \hist \in \allhist, \ar_1 \in \allargs: (\hist, \ar_0) \nileadstoe[i] ¬\phi \land (\hist, \ar_0) \ileadstoe[i] \phi \land (\hist, \ar_0, \ar_1) \nileadstoe[i] ¬\phi \land (\hist, \ar_0, \ar_1) \ileadstoe[i] \phi$.
\footnote{TODO Assume $i$, given $\ar$ and $\ar'$, picks repeatedly (in various circumstances) nonempty subsets of $\set{\prop, ¬\prop}$. Define $(\ar', \ar) \ileadstoe (\prop, \text{sure})$ iff she sometimes chooses $\set{\prop}$ and $(\ar', \ar) \ileadstoe (\prop, \text{poss})$ iff she sometimes chooses $\set{\prop}$ or $\set{\prop, ¬\prop}$. Then, $(\ar', \ar) \ileadstost (\prop, \text{sure})$ iff she always chooses $\set{\prop}$ and $(\ar', \ar) \ileadstost (\prop, \text{poss})$ iff she always chooses either $\set{\prop}$ or $\set{\prop, ¬\prop}$.}

@@

\subsection{Classify propositions}
\commentOC{Check if I current definition of $\ileadstost$ is correct, in which case I can get rid of $\ileadstostp$.}
\begin{axiom}[No contradiction]
	\label{ax:nocontr}
	$\forall \phi \in \Phi: \phi \in \iPhi ⇒ ¬\phi \notin \iPhi$.
\end{axiom}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall \prop \in \allprops$: 
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propposs] ⇒ \propposs \in \iPhi;
	\end{equation}
	and
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propsure \land \ar \ileadstostp \propposs] ⇒ \propsure \in \iPhi.
	\end{equation}
\end{axiom}

@@

For analytical clarity, it is desirable to check that \cref{ax:nocontr,ax:norm} do not hide a requirement on the content of $\ileadstostp$, thus, that they are together satisfiable without constraining $\ileadstostp$. To define this more precisely, consider temporarily these axioms as properties of a set $\iPhi \subseteq \Phi$. I write $A^1(\iPhi)$ if “axiom” 1 (interpreted as a property of $\iPhi$) holds on $\iPhi$, and similarly for $A^2(\iPhi)$. The required check amounts to prove that, given any $T, ¬, \allargs, I$, $\forall \ileadstoe: \exists \iPhi \subseteq \Phi \suchthat A^1(\iPhi) \land A^2(\iPhi)$.

Given a normative theory, define $\Box\Phi \subseteq \Phi$ as 
$\propposs \in \Box\Phi$ iff $[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propposs]$
and 
$\propsure \in \Box\Phi$ iff $[\exists \ar \in \allargs \suchthat \ar \ileadstostp \propsure \land \ar \ileadstostp \propposs]$.
Define  $P^* \subseteq \Phi$ as the set of all \emph{decidable} propositions: $P^* = \set{\phi \in \Phi \suchthat \phi \in \Box\Phi \lor ¬\phi \in \Box\Phi}$; and define $P \subseteq \Phi$ as decidable iff $P \subseteq P^*$. 
The propositions in $P^*$ are those on which the normative theory permits to take a position. It will generally be unknown in practical applications, and may be empty.

For any decidable proposition: $\phi \in \Box\Phi \oplus ¬\phi \in \Box\Phi$, where $\oplus$ denotes the exclusive disjunction; equivalently: $\phi \in \Box\Phi ⇔ ¬\phi \notin \Box\Phi$. This is proven below, and shows that \cref{ax:nocontr,ax:norm} are satisfiable together (suffices to define $\iPhi = \Box\Phi$). Also proven below is that, for that same subset of propositions: $\phi \in \iPhi \oplus ¬\phi \in \iPhi$, which ensures an unambiguous interpretation of the content of $\iPhi$ for those propositions whose status is known thanks to the model.

\begin{proposition}[Protocol coherence]
	\label{prop:protcoh}
	$[\exists \ar \in \allargs \suchthat \ar \ileadstost[pos] \phi] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost[pos] ¬\phi]$.
\end{proposition}
\begin{proof}
	Protocol coherence is satisfied by definition of $\ileadstost[pos]$. Indeed, if for some $\ar \in \allargs$, $\ar \ileadstost[pos] \phi$, then $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadstoe ¬\phi$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost[pos] ¬\phi$ is false, because it requires that $\forall \hist \in \allhist, \ar_1 \in \allargs: (\hist, \ar', \ar_1) \ileadstoe ¬\phi$, and thus in particular, that $(\ar', \ar) \ileadstoe ¬\phi$.
\end{proof}

\begin{proposition}[Internal interpretability]
	\label{th:intinterpr}
	$\forall \phi \suchthat [\phi \in \Box\Phi \lor ¬\phi \in \Box\Phi]: \phi \in \Box\Phi \oplus ¬\phi \in \Box\Phi$.
\end{proposition}
\begin{proof}
	Consider $\phi \suchthat [\phi \in \Box\Phi \lor ¬\phi \in \Box\Phi]$. Proving the claim only requires to show that $\phi \in \Box\Phi ⇒ ¬\phi \notin \Box\Phi$. By definition, $\phi \in \Box\Phi$ implies that $\exists \ar \suchthat \ar \ileadstost[pos] \phi$, and the implication is thus proven by using \cref{prop:protcoh}: $\phi \in \Box\Phi ⇒ [\exists \ar \suchthat \ar \ileadstost[pos] \phi] ⇒ ¬[\exists \ar \suchthat \ar \ileadstost[pos] ¬\phi] ⇒ \phi \notin \Box\Phi$.
\end{proof}
\begin{definition}[Restricted interpretability]
	\label{def:restrinterpr}
	$\forall \phi \suchthat [\phi \in \Box\Phi \lor ¬\phi \in \Box\Phi]: \phi \in \iPhi \oplus ¬\phi \in \iPhi$.
\end{definition}
\begin{proposition}[Restricted interpretability satisfaction]
	All models satisfying \cref{ax:nocontr,ax:norm} satisfy Restricted interpretability.
\end{proposition}
\begin{proof}
	This follows from \cref{th:intinterpr}.
\end{proof}

@@

\section{Empirical theories}
Given $\hist \in \allhist$ and $i \in I$, the claim $\hist \ileadstoe[i] \phi$ may not be verifiable, and is too specific to be falsifiable. However, given $\hist \in \allhist$, a broader claim is falsifiable: that for a given set of individuals $Q \subseteq I, \forall i \in Q, \hist \ileadstoe[i] \phi$ (provided $Q$ is big enough). Indeed, that claim it is equivalent to $\nexists i \in Q \suchthat \hist \nileadstoe[i] \phi$; exhibiting an $i \in Q$ such that $\hist \nileadstoe[i] \phi$ thus falsifies the broader claim.

\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}

An empirical theory is defined only given some normative theory. In the following, any reference to an empirical theory should always be understood as an empirical theory for some (possibly implicitly defined) normative theory.

Given $\phi \in {\gleadsto}(\allargs)$, define $\gargs[\phi, 0] \subseteq \allargs$ as $\gargs[\phi, 0] = {\gleadstoinv}(\phi)$; and, for $k \in \N$, define $\gargs[\phi, k + 1] = \gargs[\phi, k] + [{\gbeatsinv} \circ {\ibeatseinv[!\phi]}](\gargs[\phi, k]) = \set{\ar_k + \ar \suchthat \ar_k \in \gargs[\phi, k], \ar  \in [{\gbeatsinv} \circ {\ibeatseinv[!\phi]}](\gargs[\phi, k])}$.
Define $\gargs[\phi] = \bigcup_{k \in \N} \gargs[\phi, k]$, and $\gargs = \cup_{\phi \in {\gleadsto}(\allargs)} \gargs[\phi]$.

\subsection{Conditions with history}
\begin{remark}
	In fact, the model is tasked with finding a stable point after interrogating $i$. The process of trying counter-arguments $\ar_1, …$ is $q(i)$. When done, the model claims that $\ar_0 + \ar_2 + …$ is decisive. The model also claims that another process of interrogation does not lead to different answers. This is falsifiable.
	When claiming that $(\ar_1, \ar_0) \ileadstost[i] \phi$, the model really claims that $\nexists p, i' \in q^{-1}(q(i)) \suchthat p, \ar_1, \ar_0 \ileadsto[i'] ¬\phi \lor p, \ar_0, \ar_1 \ileadsto[i'] ¬\phi$, where $i' \in q^{-1}(q(i))$ iff $q(i) = q(i')$.

@@

Given $\ar_0 \in \allargs$ and $\args_2 \subseteq \allargs$, define the set $\args_1 \subseteq \allargs$ of arguments that $\args_2$ protects $\ar_0$ from as $S_1 = [f^{-1}(\args_2)]^{-1}(\ar_0) = \set{\ar_1 \in \allargs \suchthat f(\ar_1, \ar_0) \cap \args_2 ≠ \emptyset}$. 
Define $g(\ar_0) \subseteq \allargs × \allargs$ as $g(\ar_0) = \set{(\ar_2, \ar_1) \suchthat \ar_2 \in f(\ar_1, \ar_0)} = \bigcup_{\ar_1 \in \allargs} (f(\ar_1, \ar_0) × \set{\ar_1})$. The set $g(\ar_0)(\allargs)$ are the planned attacks on $\ar_0$ and the set $g(\ar_0)^{-1}(\ar_1)$ are the planned protections of $\ar_0$ against $\ar_1$. 
Then, $\args_1 = g(\ar_0)(\args_2)$.

 Define $\hist \ileadstosts \phi$ iff $[\phi = \propposs ⇒ \hist \ileadstostp \phi] \land [\phi = \propsure ⇒ \hist \ileadstostp \phi \land \hist \ileadstostp \propposs]$.

$\ar_0$ defends $\phi$ under protection of $S_2$ iff, defining $\args_1 = g(\ar_0)(\args_2)$, $\forall \hist \in \allhist: [\args_1 \cap \hist = \emptyset] ⇒ (\hist, \ar_0) \ileadstosts \phi$.

\begin{definition}[Convincingness]
	$\forall i \in I, \phi \in {\gleadsto}(\allargs), \exists k \in 2\N$ and a finite sequence $(\args_j)_{j \in 2\N, j ≤ k} \suchthat \emptyset ≠ \args_0 \subseteq {\gleadstoinv}(\phi)$ and $\forall j \in 2\N, j ≤ k, \forall \ar_j \in S_j: \ar_j$ defends $\phi$ under protection of $S_{j + 2}$ (defining $S_{k + 2} = \emptyset$) and $S_j \subseteq \cup_{\ar_{j - 2} \in \args_{j - 2}, \ar_{j - 1} \in \allargs} f(\ar_{j - 1}, \ar_{j - 2})$.
\end{definition}
Note that requiring that $\args_0$ be a singleton does not modify the condition, in the sense that any theory satisfying the condition also satisfies its strenghtening. Similarly, requiring that $S_k ≠ \emptyset$ does not modify the condition (because if some $k$ satisfy the condition with $S_k = \emptyset$, then picking $k' = k-2$ with the same sets $S_j$ also satisfy the condition, and because $S_0 ≠ \emptyset$, this recursion must end with a suitable $k$).

@@

\subsection{Checking in practice}
NOW: need to establish how convincingness can be checked in practice. This should also justify the complexity of this condition (VS simply demanding a decisive argument).

Given any $(\alltopic, ¬, \allhist, I)$, an \emph{observation protocol} (with memory) is a binary tree whose nodes are among $I × \allhist × \Phi$ and whose two edges out of any non leaf node are labeled by “yes” and “no”. The observation protocol is said to assume memory iff for each node $n = (i, \hist, \phi)$ about $i$, any node $n' = (i, {\hist}', \phi')$ about $i$ that is further down the tree continues the history, that is, the sequence ${\hist}'$ starts with $\hist$. Given an observation protocol and observables $(\ileadstoe[i])_{i \in I}$, the \emph{observed path} refers to the path in the tree defined recursively as consisting of the root node and, for each node $n = (i, \hist, \phi)$ in the path, the child node along the branch “yes” if $\hist \ileadstoe[i] \phi$ and the child node along the branch “no” otherwise. The \emph{observations} refer to the set of nodes along the observed path.

Given any $(\alltopic, ¬, \allhist, I)$, a \emph{normative claim} is any proposition involving the atoms $\iPhi$ and $\ileadstoe$ which, together with \cref{ax:nocontr,ax:norm}, permits to deduce that $\phi \in \iPhi$ for some $i \in I$ and $\phi \in \Phi$. An observation protocol, together with observables $(\ileadstoe[i])_{i \in I}$, \emph{falsifies} a normative claim iff the falsity of the claim is deducible from \cref{ax:nocontr,ax:norm} and the observations. A normative claim is \emph{falsifiable} iff there exists an observation protocol such that for any observables $(\ileadstoe[i])_{i \in I} \in (\allhist × \Phi)^I$, if the claim is false, then it is possible to deduce it from $\cref{ax:nocontr,ax:norm}$ and the observations given the observables.

\commentOC{I want: An empirical theory for a normative theory is something like a normative and falsifiable claim. It is also configurable: it must be possible to guarantee that the falsification be easy (if the claim is false) depending on the kind of claim.}

\commentOC{Falsifiable may not be the right word. It suggests that if a claim is not falsifiable, then a weaker claim is not falsifiable either, which does not hold. My falsifiability criterion demands that if a claim is false, we can see it; whether Popper’s demands that the claim be possibly false. Perhaps I should demand that we can pick a set of consequences of the claim and show it wrong it at least one possible state of the universe.}

Validity itself fails falsifiability.

>>>>

\begin{document}
\title{Empirical normative theories}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
	\href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

@@

Given $\phi \in \Phi$, define ${\ibeatse[!\phi]} \subseteq (\allhist × \allargs) × \allargs$ such that $(\hist, \ar_1) \ibeatse[!\phi] \ar_0$ iff $[(\hist \ileadstoe ¬\phi) \lor (\hist \nileadstoe \phi)] \land [\hist \text{ ends with } \set{\ar_1, \ar_0}]$.
I also write that $\ar_1 \ibeatse[!\phi] \ar_0$ iff $\exists \hist \in \allhist \suchthat (\hist, \ar_1) \ibeatse[!\phi] \ar_0$, and $\hist \ibeatse[!\phi] \ar_0$ iff $[(\hist \ileadstoe ¬\phi) \lor (\hist \nileadstoe \phi)] \land [(\exists \ar_1 \suchthat \hist \text{ ends with } \set{\ar_1, \ar_0}) \lor (\hist = \set{\ar_0})]$.

Given any $\hist \in \allhist, \ar_0 \in \allargs, \phi \in \Phi$, define $(\hist, \ar_0) \ileadstost[i] \phi$ iff $\forall \ar_1 \in \allargs: (\hist, \ar_0) \nileadstoe[i] ¬\phi \land (\hist, \ar_0) \ileadstoe[i] \phi \land (\hist, \ar_0, \ar_1) \nileadstoe[i] ¬\phi \land (\hist, \ar_0, \ar_1) \ileadstoe[i] \phi$. 
Define $(\hist, \ar_0) \ileadstosts \phi$ iff $[(\hist, \ar_0) \ileadstost \phi] \land [\phi = \propsure ⇒ (\hist, \ar_0) \ileadstost \propposs]$.
Define $\ar_0 \ileadstost[i] \phi$ iff $\forall \hist \in \allhist, (\hist, \ar_0) \ileadstost[i] \phi$, and define similarly $\ar \ileadstosts \phi$ iff $\forall \hist \in \allhist, (\hist, \ar_0) \ileadstosts \phi$.
Note in particular that: 
\begin{equation}
	\label{eq:st-in-terms-of-e}
	\ar_0 \ileadstosts \phi ⇔ [\forall \ar_1: \ar_1 \nibeatse[!\phi] \ar_0] \land [\ar_0 \nileadstoe ¬\phi] \land [\ar_0 \ileadstoe \phi].
\end{equation}
\footnote{TODO Assume $i$, given $\ar$ and $\ar'$, picks repeatedly (in various circumstances) nonempty subsets of $\set{\prop, ¬\prop}$. Define $(\ar', \ar) \ileadstoe (\prop, \text{sure})$ iff she sometimes chooses $\set{\prop}$ and $(\ar', \ar) \ileadstoe (\prop, \text{poss})$ iff she sometimes chooses $\set{\prop}$ or $\set{\prop, ¬\prop}$. Then, $(\ar', \ar) \ileadstost (\prop, \text{sure})$ iff she always chooses $\set{\prop}$ and $(\ar', \ar) \ileadstost (\prop, \text{poss})$ iff she always chooses either $\set{\prop}$ or $\set{\prop, ¬\prop}$.}

@@

\subsection{Classify propositions}
\commentOC{Check if I current definition of $\ileadstost$ is correct, in which case I can get rid of $\ileadstostp$.}
\begin{axiom}[No contradiction]
	\label{ax:nocontr}
	$\forall \phi \in \Phi: \phi \in \iPhi ⇒ ¬\phi \notin \iPhi$.
\end{axiom}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall \phi \in \Phi: 
		[\exists \ar \in \allargs \suchthat \ar \ileadstosts \phi] ⇒ \phi \in \iPhi.$
\end{axiom}

@@

For analytical clarity, it is desirable to check that \cref{ax:nocontr,ax:norm} do not hide a requirement on the content of $\ileadstost$, thus, that they are together satisfiable without constraining $\ileadstost$. To define this more precisely, consider temporarily these axioms as properties of a set $\iPhi \subseteq \Phi$. I write $A^1(\iPhi)$ if “axiom” 1 (interpreted as a property of $\iPhi$) holds on $\iPhi$, and similarly for $A^2(\iPhi)$. The required check amounts to prove that, given any $T, ¬, \allargs, I$, $\forall {\ileadstoe} \subseteq  \allhist × \Phi: \exists \iPhi \subseteq \Phi \suchthat A^1(\iPhi) \land A^2(\iPhi)$.

Given a normative theory, define $\Box\Phi \subseteq \Phi$ as 
$\Box\Phi = \set{\phi \in \Phi \suchthat \exists \ar \in \allargs \suchthat \ar \ileadstosts \phi}$.
Define  $P^* \subseteq \Phi$ as the set of all \emph{decidable} propositions: $P^* = \set{\phi \in \Phi \suchthat \phi \in \Box\Phi \lor ¬\phi \in \Box\Phi}$; and define $P \subseteq \Phi$ as decidable iff $P \subseteq P^*$. 
The propositions in $P^*$ are those on which the normative theory permits to take a position. It will generally be unknown in practical applications, and may be empty.

For any decidable proposition: $\phi \in \Box\Phi \oplus ¬\phi \in \Box\Phi$, where $\oplus$ denotes the exclusive disjunction; equivalently: $\phi \in \Box\Phi ⇔ ¬\phi \notin \Box\Phi$. This is proven below, and shows that \cref{ax:nocontr,ax:norm} are satisfiable together (suffices to define $\iPhi = \Box\Phi$). Also proven below is that, for any decidable proposition: $\phi \in \iPhi \oplus ¬\phi \in \iPhi$, which ensures an unambiguous interpretation of the content of $\iPhi$ for those propositions whose status is known thanks to the model.

\begin{proposition}[Protocol coherence]
	\label{prop:protcoh}
	$[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi]$.
\end{proposition}
\begin{proof}
	Protocol coherence is satisfied by definition of $\ileadstost$. Indeed, if for some $\ar \in \allargs$, $\ar \ileadstost \phi$, then $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadstoe ¬\phi$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost ¬\phi$ is false, because it requires that $\forall \hist \in \allhist, \ar_1 \in \allargs: (\hist, \ar', \ar_1) \ileadstoe ¬\phi$, and thus in particular, that $(\ar', \ar) \ileadstoe ¬\phi$.
\end{proof}

\begin{proposition}[Internal interpretability]
	\label{th:intinterpr}
	$\forall \phi \suchthat [\phi \in \Box\Phi \lor ¬\phi \in \Box\Phi]: \phi \in \Box\Phi \oplus ¬\phi \in \Box\Phi$.
\end{proposition}
\begin{proof}
	Consider $\phi \suchthat [\phi \in \Box\Phi \lor ¬\phi \in \Box\Phi]$. Proving the claim only requires to show that $\phi \in \Box\Phi ⇒ ¬\phi \notin \Box\Phi$. By definition, $\phi \in \Box\Phi$ implies that $\exists \ar \suchthat \ar \ileadstost \phi$, and the implication is thus proven by using \cref{prop:protcoh}: $\phi \in \Box\Phi ⇒ [\exists \ar \suchthat \ar \ileadstost \phi] ⇒ ¬[\exists \ar \suchthat \ar \ileadstost ¬\phi] ⇒ \phi \notin \Box\Phi$.
\end{proof}
\begin{definition}[Restricted interpretability]
	\label{def:restrinterpr}
	$\forall \phi \suchthat [\phi \in \Box\Phi \lor ¬\phi \in \Box\Phi]: \phi \in \iPhi \oplus ¬\phi \in \iPhi$.
\end{definition}
\begin{proposition}[Restricted interpretability satisfaction]
	All models satisfying \cref{ax:nocontr,ax:norm} satisfy Restricted interpretability.
\end{proposition}
\begin{proof}
	This follows from \cref{th:intinterpr}.
\end{proof}

@@

\section{Empirical theories}
\subsection{Falsifiability}
\commentOC{I want: An empirical theory for a normative theory is something like a normative and falsifiable claim. It is also configurable: it must be possible to guarantee that the falsification be easy (if the claim is false) depending on the kind of claim.}

Given $\hist \in \allhist$ and $i \in I$, the claim $\hist \ileadstoe[i] \phi$ may not be verifiable, and is too specific to be falsifiable. However, given $\hist \in \allhist$, a broader claim is falsifiable: that for a given set of individuals $Q \subseteq I, \forall i \in Q, \hist \ileadstoe[i] \phi$ (provided $Q$ is big enough). Indeed, that claim it is equivalent to $\nexists i \in Q \suchthat \hist \nileadstoe[i] \phi$; exhibiting an $i \in Q$ such that $\hist \nileadstoe[i] \phi$ thus falsifies the broader claim.

Given any $(\alltopic, ¬, \allargs, I)$, an \emph{observation protocol} is a binary tree whose non-leaf nodes are among $I × \allhist × \Phi$ and whose two edges out of any non leaf node are labeled by “yes” and “no”; the leaf nodes being simply $\emptyset$. The observation protocol is said to assume memory iff for each non-leaf node $n = (i, \hist, \phi)$ about $i$, any node $n' = (i, {\hist}', \phi')$ about $i$ that is further down the tree continues the history, that is, the sequence ${\hist}'$ starts with $\hist$. Given an observation protocol and observables $(\ileadstoe[i])_{i \in I}$, the \emph{observed path} refers to the path in the tree defined recursively as consisting of the root node and, for each non-leaf node $n = (i, \hist, \phi)$ in the path, the child node along the branch “yes” if $\hist \ileadstoe[i] \phi$ and the child node along the branch “no” otherwise. The \emph{observations} refer to the set of nodes along the observed path.

Given any $(\alltopic, ¬, \allargs, I)$, a \emph{normative claim} about $(\alltopic, ¬, \allargs, I)$ (or about $\allargs$, when the other defined objects are left implicit) is any first-order logic proposition involving the atoms $[(\hist, \phi) \in {\ileadstoe[i]}], \hist \in \allhist, \phi \in \Phi$ which, together with \cref{ax:nocontr,ax:norm}, permits to deduce that $\phi \in \iPhi$ for some $i \in I$ and $\phi \in \Phi$. The claim is said to be \emph{about} the given set $\allargs$. An observation protocol, together with observables $(\ileadstoe[i])_{i \in I}$, \emph{falsifies} a normative claim iff the falsity of the claim is deducible from \cref{ax:nocontr,ax:norm} and the observations. A normative claim is \emph{falsifiable} iff there exists a finite observation protocol such that for any observables $(\ileadstoe[i])_{i \in I} \in (\allhist × \Phi)^I$, if the claim is false given these observables, then it is possible to deduce it from \cref{ax:nocontr,ax:norm} and the observations given the observables.

\commentOC{Falsifiable may not be the right word. Perhaps “provable“, meaning “tautological or Popper-falsifiable”. “Falsifiable” suggests that if a claim is not falsifiable, then a weaker claim is not falsifiable either, which does not hold. My falsifiability criterion demands that if a claim is false, we can see it; whether Popper’s demands that the claim be possibly false. Perhaps I should demand that we can pick a set of consequences of the claim and show it wrong it at least one possible state of the universe.}

A normative claim is \emph{trivial} iff it is about a singleton $\allargs = \set{\ar}$, and has the form: $\ar \ileadstosts \phi$, for some $i \in I$. A normative claim is minimal iff it has the form: $\exists \ar \in \allargs \suchthat \ar \ileadstosts \phi$, for some $i \in I$ and $\phi \in \Phi$. Note that given a minimal claim, $\card{\allargs} = 1$ is equivalent to the claim being trivial.

\begin{proposition}[Trivial claims are not falsifiable]
	Given any $(\alltopic, ¬, \allargs ≠ \emptyset, I)$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Consider any claim about some $i$ and some $\phi$ and any finite observation protocol. We have to show that for some observables, the claim is false, but it is impossible to observe it (meaning, to deduce it from \cref{ax:nocontr,ax:norm} and observations).

	Given $k \in \N$, define $\ileadstosts[k]$ as $\ar^k \ileadstosts \phi$ and $\ar^{k+1} \ileadstosts ¬\phi$ (where $\ar^k$ designates a finite sequence repeating $k$ times $\ar$, for $\ar \in \allargs$). I claim that for some $k$, the observables that include this relation makes the claim false, but not observably so. This is because the observation protocol is finite, thus, only permits to observe some finite number of instances of the relation $\ileadsto$. Suffices to define $k$ as that number.
\end{proof}

There is another reason for non falsifiability, which does not involve the problem of finiteness.
\begin{proposition}[Minimal non-trivial claims are not falsifiable]
	Given any $(\alltopic, ¬, \allargs, I), \card{\allargs} ≥ 2$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Define $\ar_1$ as the first element of the sequence $\hist$ of the root node of the observation protocol (if the observation protocol is empty, suffices to define any observables that make the claim false). Define the observables so that $\forall \hist \in \allhist: \ar_2, \hist \ileadsto ¬\phi$. This proves that no argument is decisive, hence $\phi \notin \Box\Phi$, and the claim is false. But the observables may also be defined so as to satisfy the protocol (defined as: when going through the protocol, the observations can’t fail the claim).
\end{proof}

\subsection{Definition}
A general empirical theory for a normative theory $(\allprops, ¬, \allargs, I, (\ileadstoe[i])_{i \in I})$ is a tuple $({\gleadsto}, f)$, where ${\gleadsto} \subseteq \allargs × \Phi$ and $f \subseteq (\allargs × \allargs) × \allargs$ with $f(\allargs × \allargs)$ being finite.

Define $g(\ar_0) \subseteq \allargs × \allargs$ as $g(\ar_0) = \set{(\ar_2, \ar_1) \suchthat \ar_2 \in f(\ar_1, \ar_0)} = \bigcup_{\ar_1 \in \allargs} (f(\ar_1, \ar_0) × \set{\ar_1})$. The set $g(\ar_0)(\allargs)$ contains the planned attacks on $\ar_0$; the set $g(\ar_0)(\ar_2)$ contains the planned attacks on $\ar_0$ for which $\ar_2$ is a planned response; and the set $g(\ar_0)^{-1}(\ar_1) = f(\ar_1, \ar_0)$ contains the planned protections of $\ar_0$ against $\ar_1$.

Given $\ar_0 \in \allargs$ and $\args_2 \subseteq \allargs$, define the set $\args_1 \subseteq \allargs$ of arguments that $\args_2$ protects $\ar_0$ from as $\args_1 = [f^{-1}(\args_2)]^{-1}(\ar_0) = \set{\ar_1 \in \allargs \suchthat f(\ar_1, \ar_0) \cap \args_2 ≠ \emptyset} = g(\ar_0)(\args_2) = dom(f^{-1}(\args_2) \cap \allargs × \set{\ar_0})$.

\subsection{A general condition sufficient for validity and (hopefully) falsifiability}
It should be possible to define this condition as a claim that does not involve $f$, so that it can be done before defining an empirical theory. Then, show that a normative and falsifiable claim must have the form of a convincingness claim.

\begin{remark}
	In fact, the model is tasked with finding a stable point after interrogating $i$. The process of trying counter-arguments $\ar_1, …$ is $q(i)$. When done, the model claims that $\ar_0 + \ar_2 + …$ is decisive. The model also claims that another process of interrogation does not lead to different answers. This is falsifiable.
	When claiming that $(\ar_1, \ar_0) \ileadstost[i] \phi$, the model really claims that $\nexists p, i' \in q^{-1}(q(i)) \suchthat p, \ar_1, \ar_0 \ileadsto[i'] ¬\phi \lor p, \ar_0, \ar_1 \ileadsto[i'] ¬\phi$, where $i' \in q^{-1}(q(i))$ iff $q(i) = q(i')$.

@@

$\ar_0$ $i$-defends $\phi$ when ignoring $S_1$ iff $\forall \hist \in \allhist: [\args_1 \cap \hist = \emptyset] ⇒ (\hist, \ar_0) \ileadstosts \phi$.
$\ar_0$ $i$-defends $\phi$ under protection of $S_2$ iff $\ar_0$ $i$-defends $\phi$ when ignoring $g(\ar_0)(\args_2)$.

\begin{definition}[Convincingness]
	$\forall i \in I, \phi \in {\gleadsto}(\allargs), \exists k \in 2\N$ and a finite sequence $(\args_j)_{j \in 2\N, j ≤ k} \suchthat \emptyset ≠ \args_0 \subseteq {\gleadstoinv}(\phi)$ and $\forall j \in 2\N, j ≤ k, \forall \ar_j \in S_j: \ar_j$ $i$-defends $\phi$ under protection of $S_{j + 2}$ (defining $S_{k + 2} = \emptyset$) and $S_j \subseteq \cup_{\ar_{j - 2} \in \args_{j - 2}, \ar_{j - 1} \in \allargs} f(\ar_{j - 1}, \ar_{j - 2})$.
\end{definition}
Note that requiring that $\args_0$ be a singleton does not modify the condition, in the sense that any theory satisfying the condition also satisfies its strenghtening. Similarly, requiring that $S_k ≠ \emptyset$ does not modify the condition (because if some $k$ satisfy the condition with $S_k = \emptyset$, then picking $k' = k-2$ with the same sets $S_j$ also satisfy the condition, and because $S_0 ≠ \emptyset$, this recursion must end with a suitable $k$).

@@

\begin{theorem}
	The convincingness claim of any empirical theory is falsifiable: given an ET, there exists a finite obs prot with memory such that for any observables, if the claim is false given these observables, then it is possible to deduce it from the axioms and observations.
\end{theorem}
\begin{proof}

\end{proof}

\section{Additive empirical theories}
TODO: 3) define single-answer models. 4) phrase conditions below for history and single-answer models. 5) show that these conditions can be verified or falsified. 6) show that if the claim of the model does not include these conditions, then the model is not falsifiable, as suggested here below for some form of claims.

\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
${\gbeatsinv}(\allargs)$ finite,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}
\commentOC{${\gbeatsinv}$ need not be finite, only those even ones: the transitive closure of ${\gleadstoinv}(\Phi)$ through ${\gbeatsinv}^2$, that is, $({\gbeatsinv})^2[{\gleadstoinv}(\Phi)] = \cup_{k \in 2\N}({\gbeatsinv}^{k}({\gleadstoinv}(\Phi)))$.}
\commentOC{I should only define $\ar \in \allargs$ plus $\ard$.}

This defines a function: $\forall \ar_1 \gbeats \ar_0: f(\ar_1, \ar_0) = \ar_0 + {\gbeatsinv}(\ar_1) = \set{\ar_0 + \ard_2 \suchthat \ard_2 \in {\gbeatsinv}(\ar_1)}$, and $f(\ar_1, \ar_0) = \emptyset$ otherwise. Thus, $f = \set{((\ar_1, \ar_0), \ar_2) \suchthat \exists \ard_2 \in \allargs \suchthat \ard_2 \gbeats \ar_1 \gbeats \ar_0 \land \ar_2 = \ar_0 + \ard_2}$.
Also, from the previous definition, $g(\ar_0)(\ar_0 + \argsd_2) = dom(f^{-1}(\ar_0 + \argsd_2) \cap (\allargs × \set{\ar_0})) = {\gbeats}(\argsd_2) \cap {\gbeatsinv}(\ar_0)$ because $(\ar_1, \ar_0) \in f^{-1}(\ar_0 + \argsd_2) ⇔ \exists \ar_2 \in \ar_0 + \argsd_2 \suchthat \ar_2 \in f(\ar_1, \ar_0) ⇔ \ar_1 \gbeats \ar_0 \land \exists \ard \in \argsd_2 \suchthat \ard \gbeats \ar_1$.
\commentOC{Could be false. Define }
